\scshape Topics in mechanism design

### 2025-12-07 ch1.tex
Add files via upload

Note that $\mathcal{Q}$ is a convex space: if $q$ and $q'$ are both increasing maps $[0,1] \to [0,1]$, then so is $\alpha q + (1-\alpha) q'$ for any scalar $\alpha \in [0,1]$.
Note further that the objective $R$ is linear:
%
\begin{equation*}
	R( \alpha q + (1-\alpha) q' ) = \alpha R(q) + (1-\alpha) R(q') 
	\quad \text{for any $q,q' \in \mathcal{Q}$ and $\alpha \in [0,1]$.}
\end{equation*}
%
$R$ is also continuous in a suitable sense.
Finally, $\mathcal{Q}$ is (in fact) compact in a suitable sense.

@@

		\item
		Suppose to begin with that the monopolist is a `maxmax' decision-maker, meaning that she evaluates an allocation $q$ according to the most optimistic of the beliefs $F$ in the set $\mathcal{F}$:%
			\footnote{This could be interpreted as `motivated reasoning'.}
		%
		\begin{equation*}
			\mathcal{U}(q) = \max_{F \in \mathcal{F}} R_F(q) .
		\end{equation*}
		%
		Prove that there is a posted-price mechanism that is optimal.
		(Hint: $\mathcal{U}$ is not linear. But\dots?)

@@

This intuition is incomplete, it seems to me, at least when $\inf S > 0$, because `perfect competition' surely means selling at marginal cost (which we've normalised to zero), rather than at price $\inf S$!
Here's a better intuition that I learned from Francesco Nava: the seller's lack of commitment constrains her eventually to sell the good whatever the buyer's value, and constrains her no further than that; thus she sells the good at the highest market-clearing price, which is $\inf S$.

@@

Let's write $\mathcal{Q}_2$ for all cyclically monotone allocations.
$\mathcal{Q}_2$ is convex (and compact),
and revenue is a linear function $R : \mathcal{Q}_2 \to \R$,
so our previous analysis tells us that there must be an optimal mechanism that is an extreme point of $\mathcal{Q}_2$.
But characterising these extreme points is hard, and there are many of them.
More broadly: cyclic monotonicity is intractable.

@@

In the model of this chapter, the payoff-relevant ex-post outcomes are $(g,p) \in \{0,1\} \times \R$, where $g=1$ iff the agent gets the good.
A \emph{lottery} (over outcomes) is a probability measure on $\{0,1\} \times \R$.
Each type of the agent has a preference over lotteries.
\emph{Expected-utility (EU) preferences} over lotteries $\ell$ are those that admit a \emph{linear} representation:
$U(\ell) = \int_{\{0,1\} \times \R} u \dd \ell$
for some (`utility index') $u : \{0,1\} \times \R \to \R$.
(I have omitted the measure-theoretic details.)
You will recall that $u$ is unique up to affine transformations:
if (and only if) $U(\ell) = \int_{\{0,1\} \times \R} u \dd \ell$
and $V(\ell) = \int_{\{0,1\} \times \R} v \dd \ell$ represent the same EU preference over lotteries, then $u = \alpha + \beta v$ for some $\alpha \in \R$ and $\beta>0$.

>>>>

Note that $\mathcal{Q}$ is a convex space: if $q$ and $q'$ are both increasing maps $[0,1] \to [0,1]$, then so is $\alpha q + (1-\alpha) q'$ for any scalar $\alpha \in [0,1]$.
Note further that the objective $R$ is affine:
%
\begin{equation*}
	R( \alpha q + (1-\alpha) q' ) = \alpha R(q) + (1-\alpha) R(q') 
	\quad \text{for any $q,q' \in \mathcal{Q}$ and $\alpha \in [0,1]$.}
\end{equation*}
%
$R$ is also continuous in a suitable sense.
Finally, $\mathcal{Q}$ is (in fact) compact in a suitable sense.

@@

		\item
		Suppose to begin with that the monopolist is a `maxmax' decision-maker, meaning that she evaluates an allocation $q$ according to the most optimistic of the beliefs $F$ in the set $\mathcal{F}$:%
			\footnote{This could be interpreted as `motivated reasoning'.}
		%
		\begin{equation*}
			\mathcal{U}(q) = \max_{F \in \mathcal{F}} R_F(q) .
		\end{equation*}
		%
		Prove that there is a posted-price mechanism that is optimal.
		(Hint: $\mathcal{U}$ is not affine. But\dots?)

@@

This intuition is incomplete, it seems to me, at least when $\inf S > 0$, because `perfect competition' surely means selling at marginal cost (which we've normalised to zero), rather than at price $\inf S$!
Here's a better intuition that I learned from Francesco Nava: the seller's lack of commitment constrains her eventually to sell the good whatever the buyer's valuation, and constrains her no further than that; thus she sells the good at the highest market-clearing price, which is $\inf S$.

@@

Let's write $\mathcal{Q}_2$ for all cyclically monotone allocations.
$\mathcal{Q}_2$ is convex (and compact),
and revenue is an affine function $R : \mathcal{Q}_2 \to \R$,
so our previous analysis tells us that there must be an optimal mechanism that is an extreme point of $\mathcal{Q}_2$.
But characterising these extreme points is hard, and there are many of them.
More broadly: cyclic monotonicity is intractable.

@@

In the model of this chapter, the payoff-relevant ex-post outcomes are $(g,p) \in \{0,1\} \times \R$, where $g=1$ iff the agent gets the good.
A \emph{lottery} (over outcomes) is a probability measure on $\{0,1\} \times \R$.
Each type of the agent has a preference over lotteries.
\emph{Expected-utility (EU) preferences} over lotteries $\ell$ are those that admit an \emph{affine} representation:
$U(\ell) = \int_{\{0,1\} \times \R} u \dd \ell$
for some (`utility index') $u : \{0,1\} \times \R \to \R$.
(I have omitted the measure-theoretic details.)
You will recall that $u$ is unique up to positive affine transformations:
if (and only if) $U(\ell) = \int_{\{0,1\} \times \R} u \dd \ell$
and $V(\ell) = \int_{\{0,1\} \times \R} v \dd \ell$ represent the same EU preference over lotteries, then $u = \alpha v + \beta$ for some $\alpha > 0$ and $\beta \in \R$.
